---
title: "EMA Actigraphy merge"
output:
  html_notebook: default
  pdf_document: default
---

The goal is to derive a tidy EMA data-set including a data dictionary for CHARMS (and SPEAK) participants which is merged with the GENEActiv GGIR accelerometry (sleep) data. 

I have copied the raw EMA data (completely untouched) from /Users/jessicahartmann/OneDrive/Data and files from Orygen server june 2021/CHARMS/EMA data/EXTRACTED DATA into the current home directory 


### Getting data into  R



Load required packages

```{r}
rm(list = ls()) ##new empty work space
library(plyr)
library(readr)
library(dplyr)
library(readxl)
library(tidyr)
```

Read files into a list, then read in the list
```{r}
myfiles = list.files(path = "extracted_data", pattern="*.csv", full.names=TRUE)
ema_df <- ldply(myfiles, read.csv)
```

### Some tidying 

Keep only surveys which have answers and only items which were part of the CHARMS survey
```{r}

ema_df <- ema_df %>%
        filter (HAS_ANSWERS == "YES") %>%
        subset(select = -(NAP.1:SUBST.10_RT.1))

```


Make overview of how many surveys have been answered by participants (note: there are some test subjects still in there which will be removed later)
```{r}
ema_n <- ema_df %>% 
        group_by(PARTICIPANT_ID) %>%
        summarise(n = n()) 
```

The EMA dataset comes with its own participant id's. Now we need to add the "official" CHARMS ID which is included in Codebook_ID. 
Also added in this step: age and sex. 
```{r}
ids <- read_excel("Codebook_IDs.xlsx")
ema_df<- merge(ema_df, ids, by = "PARTICIPANT_ID") ##check if gender and age is added


ema_m <- ema_df %>%
        subset(!is.na("CHARMS_ID")) %>% ## only keep the observations which have a CHARMS ID
        select(CHARMS_ID, Gender, Age, everything())
      
```

```{r}
#write.csv(ema_m, "ema_merged.csv") ##for Holly 
```


### How many unique people are in this EMA data set?

```{r}
idlist <- unique(ema_m$CHARMS_ID) ##make a vector 
length(idlist) ##length of the vector 
```

OK, we have N = 14 people in the EMA data. 

### Aggregate EMA measures at day-level 

For Holly's project, we will aggregate the dataset to one EMA observation per person per day. This makes it easier to work with and understand the data (though losing some power)  

 Convert to **date** 
```{r}
ema_m$COMPLETED <- as.Date(ema_m$COMPLETED, format = "%d/%m/%Y")

## there are a few observations with invalid date, exclude these. Check with Ann Ee why 
ema_m <- ema_m[ which(ema_m$COMPLETED > 01/01/1970),] 
```

Create data frame with only one observation per individual per day and keep only relevant variables. This can then be merged with the sleep day summary data. 
```{r}

ema_perday <-  ema_m %>% 
            group_by(CHARMS_ID, COMPLETED) %>% 
            select (-c(ITERATION, PROGRAM_VERSION, SURVEY, HAS_ANSWERS, 
                              DELIVERED, EXPIRED, UPLOADED)) %>% ## exclude non-relevant columns
            summarise(across(where(is.numeric), list(mean), na.rm = TRUE)) %>%## calculate the mean
           rename_with(~ sub("_1$", "", .x), everything())
            
```
```{r}
##rename charms id and date for merging 
ema_perday <- rename(ema_perday, ID = CHARMS_ID, calendar_date = COMPLETED) 

```

## Calculating positive affect (PA) and negative affect (NA) 


**PA**: As per Ann Ee's thesis, calculate PA out of the following items: Cheerful, enthousiastic, down (reverse coded). 
**NA**: As per Ann Ee's thesis, calculate NA out of the following items: lonely, insecure, anxious, guilty, relaxed (reverse coded)


```{r}
## reverse score down 

ema_perday <- ema_perday %>%
              mutate(DOWNrev = 101 - DOWN) %>% ## reverse score down 
              mutate(RELAXrev = 101 - RELAXED) %>%
              mutate(pa = (CHEERF + ENTHOUS + DOWNrev)/3) %>% ## calculate PA
              mutate(na = (LONELY + INSECURE + ANXIOUS + GUILTY + 
                             RELAXrev)/5) ## calculate NA

```


## sleep data 

pre-processing: combining "day"GGIR files and "night" GGIR files to obtain all relevant variables 


Read in sleep data 
In GGIR 2.4.2, there is a problem with the id variable in the day summaries (wrong). Therefore, make new ID name based on file name 
```{r}
ggir_p4 <- read.csv("../GGIR/McKenzie_results/output_McKenzie_data/results/part4_nightsummary_sleep_cleaned.csv")
ggir_p4 <- ggir_p4 %>%
          mutate(calendar_date =  as.Date(calendar_date, "%d/%m/%Y"))

ggir_p5 <- read.csv("../GGIR/McKenzie_results/output_McKenzie_data/results/part5_daysummary_WW_L30M100V400_T5A5.csv")

ggir_p5 <- ggir_p5 %>%
          select(-ID) %>%
          mutate(id = substr(filename, 1, 4)) %>%
          mutate(ID = as.integer(id)) %>%
          mutate(calendar_date =  as.Date(calendar_date)) %>%
          select(ID, calendar_date, sleep_efficiency)

ggir <-
    dplyr::full_join(ggir_p4, ggir_p5, by = c("ID", "calendar_date"))


```

Merge data sets 

```{r}
ema_sleep_all <- merge(ema_perday, ggir, all = TRUE) ## all observations, not just ## matched ones


ema_sleep_all <- dplyr::full_join(ema_perday, ggir, by = c("ID", "calendar_date"))                                                     
```

Sort and order 

```{r}
ema_sleep_all <- ema_sleep_all %>%
                select(-sleepefficiency) %>% ##from GGIR part 4 which is missing, keep from part5
                relocate(ID, calendar_date, Age, Gender, pa, na, sleeponset:sleep_efficiency)

ema_sleep_all <- ema_sleep_all[!is.na(ema_sleep_all$ID),]  ##exclude obs without ID  

```


Create a day index 

```{r}
ema_sleep_all <- ema_sleep_all %>%
                  group_by(ID) %>%  
                  arrange(calendar_date) %>%
                  complete(ID, calendar_date = seq.Date(min(calendar_date), 
                                                       max(calendar_date), 
                                                       by = "day")) %>%
                  mutate(days = row_number()) %>% 
                  relocate(ID, calendar_date, days) #%>% 
                  #subset((!is.na(pa) & !is.na(sleeponset)))

              
  
                

```

save to send to Holly 

```{r}
write.csv(ema_sleep_all, "ema_sleep_merged.csv")
```

```{r}

```


